---
title: "RepData_Assessment2"
author: "AllenTsai"
date: "2015年9月27日"
output: html_document
---

Title: Your document should have a title that briefly summarizes your data analysis

Synopsis: Immediately after the title, there should be a synopsis which describes and summarizes your analysis in at most 10 complete sentences.

There should be a section titled  which describes (in words and code) how the data were loaded into R and processed for analysis. In particular, your analysis must start from the raw CSV file containing the data. You cannot do any preprocessing outside the document. If preprocessing is time-consuming you may consider using the cache = TRUE option for certain code chunks.

There should be a section titled Results in which your results are presented.

### Data Processing
1. use fread to read data, extract csv from bz2 file before processing

```{r, cache=TRUE, result="hide"  readdata}
setwd("/Users/hadoop")
#If you want to use fread, please install the dev version
source("RepData_PeerAssessment2/installPackage.R")

data.file <-"repdata-data-StormData.csv"
if(require(data.table) && packageVersion("data.table") >= "1.9.5" ) {
        topData <- read.table(data.file, sep=",", nrows = 50, header=TRUE)
        col.classes <- sapply(topData, class)
        raw.data <- fread(data.file, header= TRUE, sep=",", colClasses = col.classes)
} else {
        raw.data <- read.table(data.file, sep=",", header = TRUE)
}
```

2. cleaning data 
  - uppercase EVTYPE, TIME_ZONE, PROPDMGEXP, CROPDMGEXP
  - ignore PROPDMGEXP/CROPDMGEXP not in (B,M,K)
  - adjust PROPDMG by PROPDMGEXP (K = 1e3, M = 1e6, B= 1e9)
  - adjust CROPDMG by CROPDMGEXP (K = 1e3, M = 1e6, B= 1e9)
```{r cleanData, cache=TRUE, result="hide"}
cleanData <- function(data) {
        clean.data <- data %>%
                mutate(EVTYPE = gsub("[^(a-zA-Z0-9/,)]",".", toupper(EVTYPE)) ) %>%
                mutate(TIME_ZONE = toupper(TIME_ZONE)) %>%
                mutate(PROPDMGEXP = toupper(PROPDMGEXP)) %>%
                mutate(CROPDMGEXP = toupper(CROPDMGEXP)) %>%
                filter(PROPDMGEXP %in% c("B","M","K")) %>%
                filter(CROPDMGEXP %in% c("B","M","K")) %>%
                mutate(BGN_DATE = as.Date(BGN_DATE, format="%m/%d/%Y")) %>%
                mutate(BGN_YR = year(BGN_DATE)) %>%
                mutate(BGN_DECADE = cut(BGN_YR, breaks = c(1949, 1959, 1969, 1979, 1989, 1999,2012) ) )%>%
                mutate(PROPDMG = ifelse(PROPDMGEXP == "K", PROPDMG * 1000,
                                        ifelse(PROPDMGEXP == "M", PROPDMG * 1000000, 
                                               ifelse(PROPDMGEXP == "B", PROPDMG * 1000000000, PROPDMG ) ) )) %>%
                mutate(CROPDMG = ifelse(CROPDMGEXP == "K", CROPDMG * 1000, 
                                        ifelse(CROPDMGEXP == "M", CROPDMG * 1000000, 
                                               ifelse(CROPDMGEXP == "B", CROPDMG * 1000000000, CROPDMG ) ) )) 
        levels(clean.data$BGN_DECADE) <- c('1950-1959', '1960-1969','1970-1979','1980-1989','1990-1999','2000-2011')
        return (clean.data)
}


clean.data <- cleanData(raw.data)
```
3. summary clean data by Event Type
  - sum.FATALITIES
  - sum.INJURIES
  - sum.PROPDMG
  - sum.CROPDMG
  - sum.DMG = sum.PROPDMG + sum.CROPDMG
  - sum.casualties = sum.FATALITIES  + sum.INJURIES
  - index.casualties = sum.FATALITIES * 3 + sum.INJURIES
  I gave FATALITIES a factor of 3 to calculate index.casualties 
```{r summaryEventData, cache=TRUE, result="hide"}
summaryEventData <- function(cleanData) {
        summary.data <- cleanData %>%
                group_by(EVTYPE) %>%
                summarise(event.count=n() ,
                          sum.FATALITIES = sum(FATALITIES, na.rm= TRUE),
                          sum.INJURIES = sum(INJURIES, na.rm= TRUE),
                          sum.PROPDMG = sum(PROPDMG, na.rm= TRUE),
                          sum.CROPDMG = sum(CROPDMG, na.rm= TRUE)) %>%
                mutate(sum.DMG = sum.PROPDMG + sum.CROPDMG)  %>%
                mutate(sum.casualties = sum.FATALITIES  + sum.INJURIES) %>%
                mutate(index.casualties = sum.FATALITIES * 3 + sum.INJURIES)
        return (summary.data)
}  

summary.event.data <- summaryEventData(clean.data)
```

### Result
1. arrange summary.event.data order by desc(sum.DMG), get top 20 event
2. economic damaage(property and crop) by event type
```{r summary.dmg.data}
summary.dmg.data <- summary.event.data %>%
        arrange(desc(sum.DMG)) %>%
        top_n(20) %>%
        select(EVTYPE, c(1,5,6,7))

p1 <- ggplot(summary.dmg.data, aes(x=EVTYPE, y=sum.DMG))  +
        geom_bar(stat="identity")  +
        labs(title = "Total damages casuse by Top 20 Weather Event") +
        labs(y = "Estimates of Damage (property + crop) ") +
        labs(x = "Event Type") +
        theme(axis.text.x= element_text(angle=90, vjust=0.6, size=8))+
        theme(title= element_text(vjust=0.6, size=8))
```

1. order summary.event.data by desc(index.casualties), get top 20 event
2. human health impact(FATALITIES + INJURIES) by event type
```{r summary.casualties.data}
summary.casualties.data <- summary.event.data %>%
        arrange(desc(index.casualties)) %>%
        top_n(20) %>%
        select(EVTYPE, c(1,3,4,8,9))
p2 <- ggplot(summary.casualties.data, aes(x=EVTYPE, y=sum.casualties))  +
        geom_bar(stat="identity")  +
        labs(title = "Total Casualties casuse by Top 20 Weather Event") +
        labs(y = "Casualties (fatalities + injuries) ") +
        labs(x = "Event Type") +
        theme(axis.text.x= element_text(angle=90, vjust=0.6, size=8))+
        theme(title= element_text(vjust=0.6, size=8))
```

```{r}
setwd("/Users/hadoop/RepData_PeerAssessment2")
source("multiplot.R")
multiplot(p1, p2, cols=2)
```



```{r summary.damage}
summary.damage <- clean.data %>%
            filter(EVTYPE %in% as.data.frame(summary.dmg.data)[1:3,1]) %>%
            filter(TIME_ZONE %in% c("PST","MST","EST","CST")) %>%
            group_by(EVTYPE,BGN_YR,TIME_ZONE) %>%
            summarise(event.count=n() ,
                  sum.PROPDMG = sum(PROPDMG, na.rm= TRUE),
                  mean.PROPDMG = mean(PROPDMG, na.rm= TRUE),
                  sum.CROPDMG = sum(CROPDMG, na.rm= TRUE),
                  mean.CROPDMG = mean(CROPDMG, na.rm= TRUE)) %>%
        mutate(sum.Damage = sum.PROPDMG + sum.CROPDMG)  %>%
        mutate(mean.Damage = mean.PROPDMG + mean.CROPDMG) 

pd1 <-  ggplot(aes(x=BGN_YR, col=EVTYPE), data=summary.damage) +        
        geom_line(aes(y= sum.Damage))+
        facet_wrap(~ TIME_ZONE, ncol=2) +  
        labs(title = "Total damages by Top3 Event Across the United States" ) + 
        labs(y = "Estimates of Damage (property + crop)") +
        labs(x = "Year of Event")  +
        theme(title=element_text(vjust=0.5,hjust=0.5, size=7))+ 
        theme(axis.text.x= element_text(angle=0, vjust=0.5, size=8))    
plot(pd1)
```


```{r}
summary.casualties <- clean.data %>%
        filter(EVTYPE %in% as.data.frame(summary.casualties.data)[1:3,1]) %>%
        filter(TIME_ZONE %in% c("PST","MST","EST","CST")) %>%
        group_by(EVTYPE,BGN_YR,TIME_ZONE) %>%
        summarise(event.count=n() ,
                  sum.FATALITIES = sum(FATALITIES, na.rm= TRUE),
                  sum.INJURIES = sum(INJURIES, na.rm= TRUE)) %>%
        mutate(sum.Casualties = sum.FATALITIES + sum.INJURIES)  


pc1 <- ggplot(aes(x=BGN_YR, col=EVTYPE), data=summary.casualties) +        
        geom_line(aes(y= sum.Casualties, col=EVTYPE))+
        scale_x_continuous(breaks=unique(summary.casualties$BGN_YR)) +
        facet_wrap(~ TIME_ZONE) +  
        labs(title = "Total Casualties by Top3 Event Across the United States" ) + 
        labs(y = "Casualties(Fatalities + Injuries)") +
        labs(x = "Year of Event")  +
        theme(axis.text.y=element_text(vjust=0.5,hjust=0.5, size=8))+ 
        theme(axis.text.x= element_text(angle=90, vjust=0.5, size=8))    

plot(pc1)
```